Cybersecurity Impacts of the Covid-19 Pandemic in Italy

### 2021-03-09 article.tex
Small final edits

The chart in Fig.~\ref{fig:2} analyzes the trend of cyberattaks.
The figure shows that, from 2007 to 2015, in average, about 40\% of the respondents
reported cyberattacks (leftmost full red line). Since 2015, the trend increased to 45\%
of the respondents (rightmost full red line). 
In 2018, OAD reached the peak of reported attacks and, for the first time, the percent of
reported attacks became larger than that percent of companies that did not report any attack.
After the peak 
in 2018, year 2019 has a lower figure and the same occurs in the first quarter 2020.
The trend for full 2020 will be discovered with the currently ongoing OAD survey.
This view, over several years, does not show significant changes, not even inside
the pandemic period, but shows only a slight increase in cyberattacks 
in the recent years, not necessarily attributable to the pandemic crisis. One could wonder
why the percentages in Fig.~\ref{fig:2} are lower than what the reader would have expected, 
given the numerous cyber attacks occured in Italy. As already discussed at the beginning of 
this section, the main reason can be attributed to the sizes of italian companies.

The comparison between the whole 2019 and the first quarter 2020
in Fig.~\ref{fig:3} maintains the same logical trend. The only meaningful difference
is in the number of attacks reported by the smallest
organizations (up to ten employees): they were 0\% in 2019 and grow up to
22,2\% at the beginning of 2020 (see the red arrow in Fig.~\ref{fig:3}).
Such a strong increase follows from the extensive use of e-banking and 
e-commerce services, with the related cyberattacks.

@@

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{pictures/fig5.png}
  \caption{Widespread of the digital attacks on the respondents' computer systems, classified wrt.\ the type of attack, between 2019 and the first quarter of 2020 (multiple answers)~\cite{oad20}. (*) Non all the IT systems of the respondents utilize these types of solutions/systems}
  \label{fig:5}
\end{figure}

>>>>

The chart in Fig.~\ref{fig:2} analyzes the trend of cyberattaks.
The figure shows that, from 2007 to 2015, in average, about 40\% of the respondents
reported cyberattacks (leftmost full red line). Since 2015, the trend increased to 45\%
of the respondents (rightmost full red line). 
In 2018, OAD reached the peak of reported attacks and, for the first time, the percent of
reported attacks became larger than that percent of companies that did not report any attack.
After the peak 
in 2018, year 2019 has a lower figure and the same occurs in the first quarter of 2020.
The trend for full 2020 will be discovered with the currently ongoing OAD survey.
This view, over several years, does not show significant changes, not even inside
the pandemic period, but shows only a slight increase in cyberattacks 
in the recent years, not necessarily attributable to the pandemic crisis. One could wonder
why the percentages in Fig.~\ref{fig:2} are lower than what the reader would have expected, 
given the numerous cyber attacks occured in Italy. As already discussed at the beginning of 
this section, the main reason seems the size of the Italian companies.

The comparison between the whole 2019 and the first quarter of 2020
in Fig.~\ref{fig:3} maintains the same logical trend. The only meaningful difference
is in the number of attacks reported by the smallest
organizations (up to ten employees): they were 0\% in 2019 and grow up to
22,2\% at the beginning of 2020 (see the red arrow in Fig.~\ref{fig:3}).
Such a strong increase follows from the extensive use of e-banking and 
e-commerce services, with the related cyberattacks.

@@

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{pictures/fig5.png}
  \caption{Widespread of the digital attacks on the respondents' computer systems, classified wrt.\ the type of attack, between 2019 and the first quarter of 2020 (multiple answers)~\cite{oad20}. Where marked with (*), we note that not all the IT systems of the respondents use these types of solutions/systems.}
  \label{fig:5}
\end{figure}
